---
title: "STAT 240 Discussion 10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
library(tidyverse)
library(lubridate)
source("ggprob.R")
```

## Group XXX

## Members Present

- ADD THE NAMES OF EVERYONE PRESENT
- MAKE THE NAME **BOLD** OF THE RESPONSIBLE PERSON TO UPLOAD THE FILES FOR THE GROUP

## Members Absent

- ADD NAMES OF ANY ABSENT MEMBERS

## Data

The following code reads in the official Madison weather data
and calculates the average winter temperature (November through February) for each winter and graphs the data and a fitted linear model.

The official weather record in this data set extends from January 1, 1869 through December 31, 2020.
The complete winters begin with 1869-1870 in November and extend through the winter of 2019-2020 as this data set does not include the most recent winter.

```{r read-data}
## Read and transform the Madison weather data
mw_orig = read_csv("madison-weather-official-2020.csv")

mw = mw_orig %>% 
  mutate(year = year(date),
         month = month(date, label = TRUE),
         .after = date) %>% 
  mutate(year1 = case_when(
    month > "Jun" ~ year,
    month < "Jul" ~ year - 1),
    year2 = year1 + 1,
    winter = str_c(year1, "-", year2),
    .before = year) 

## add the winter variables and filter
mw_winter = mw %>%
  filter(month %in% c("Nov", "Dec", "Jan", "Feb")) %>%
  select(-prcp, -contains("snow")) %>% 
  drop_na() %>% 
  filter(winter >= "1869-1870" & winter <= "2019-2020") %>% 
  group_by(winter, year1) %>%
  summarize(tavg = mean(tavg))
```

## Setup

### Scatter plot and regression line

```{r plot-data}
ggplot(mw_winter, aes(x = year1, y = tavg)) +
  geom_point() +
  geom_smooth(method = "lm", se=FALSE) +
  xlab("Year (first year of the winter)") +
  ylab("Average Winter Temperature") +
  ggtitle("Madison, WI Average Winter Temperature",
          subtitle = "Winter defined as Nov. through Feb.") +
  theme_bw()
```

### Regression Model

The next block of code fits a linear model and then adds the residuals and predicted values as new columns for plotting purposes later on.

```{r fit-regression-model}
## fit the regression model
winter_lm = lm(tavg ~ year1, data = mw_winter)

## get summary table of coefficients and inference values
summary(winter_lm)

## augment data frame with  residuals and fitted values for plotting later
mw_winter_aug = mw_winter %>% mutate()
mw_winter_aug$.resid = resid(winter_lm)
mw_winter_aug$.pred = fitted(winter_lm)

head(mw_winter_aug)
```

<br/><br/>

## Problems


### 1 

Using the results of the linear model fit above, write out the estimated linear model by replacing the (a) and (b) with the estimated coefficients; y_hat represents the estimated average winter temperature in Madison for year x.  What is the interpretation of the estimated slope?

y_hat = (a) + (b) * x

REPLACE THIS TEXT WITH YOUR RESPONSE



### 2

### 2A

Calculate the end points of a 95% confidence interval for the slope of the regressions model.
Note that we have data for this entire period of time,
so a confidence interval here represents the slope in a linear unseen climate trend whereas the data is affected by the randomness of weather.

Round the margin of error to two significant digits and then round the estimated slope to the same accuracy.
The margin of error is the product of a quantile from a t distribution with an appropriate number of degrees of freedom.

```{r}
df <- nrow(mw_winter_aug) - length(coefficients(winter_lm))
t_quantile <- qt(0.975, df)
se_slope <- summary(winter_lm)$coefficients["year1", "Std. Error"]
estimated_slope <- summary(winter_lm)$coefficients["year1", "Estimate"]
margin_of_error <- t_quantile * se_slope
margin_of_error_rounded <- round(margin_of_error, 2)
estimated_slope_rounded <- round(estimated_slope, 2)
lower_endpoint <- estimated_slope - margin_of_error
upper_endpoint <- estimated_slope + margin_of_error
lower_endpoint_rounded <- round(lower_endpoint, 2)
upper_endpoint_rounded <- round(upper_endpoint, 2)
confidence_interval <- c(lower_endpoint_rounded, upper_endpoint_rounded)
confidence_interval
```


### 2B

Interpret this confidence interval in context, following examples from lecture.

The 95% confidence interval for the slope of the regression model is (-0.03, 0.00). We can be 95% confident that the true unseen climate trend in Madison's average winter temperature shows no significant overall increase or decrease. The interval includes zero, indicating no strong evidence for a consistent linear change over the years, accounting for weather randomness and natural variability.


### 3

Is there strong evidence that the unknown slope of the regression line of average winter temperature in Madison due to a changing climate is different from zero?

### 3A

Report a t-test statistic and p-value from the model summary output.

REPLACE THIS TEXT WITH YOUR RESPONSE


### 3B

Verify the calculation of the p-value using the `pt()` function.

```{r}

```


### 3C

Interpret this hypothesis test in the context of the question.

REPLACE THIS TEXT WITH YOUR RESPONSE


### 4

Make a density plot of the residuals from the model. Calculate the mean and standard deviation of the residuals. Overlay a normal density with these values for the mean and standard deviation.

Does it appear that a normal distribution is a reasonable approximation of the distribution of variation of points around the regression line?

```{r}

```

REPLACE THIS TEXT WITH YOUR RESPONSE
