---
title: "STAT 240 Discussion 5"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
library(tidyverse)
source("ggprob.R")
```

## Group XXX 

## Members Present

- ADD THE NAMES OF EVERYONE PRESENT
- MAKE THE NAME **BOLD** OF THE RESPONSIBLE PERSON TO UPLOAD THE FILES FOR THE GROUP

## Members Absent

- ADD NAMES OF ANY ABSENT MEMBERS

## Questions


### 1

If $X \sim \text{N}(20, 3)$, what is $P(15 < X < 27)$?

```{r}
# insert code here
mu <- 20
sigma <- 3

prob <- pnorm(27, mean = mu, sd = sigma) - pnorm(15, mean = mu, sd = sigma)

prob
```


### 2

Draw a graph of a normal distribution with mean 20 and standard deviation 3 with the area between 15 and 27 shaded.
How does this area correspond to the previous calculated probability?

> REPLACE THIS TEXT WITH YOUR RESPONSE



### 3

Define $z_1 = (15 - 20)/3$ and $z_2 = (27 - 20)/3$.
If $Z \sim \text{N}(0,1)$, what is $P(z_1 < Z < z_2)$?

```{r}
# insert code here
z1 <- (15 - 20) / 3
z2 <- (27 - 20) / 3

prob <- pnorm(z2) - pnorm(z1)

prob
```



### 4

Draw a graph of a standard normal distribution and shade the area under the curve between $z_1$ and $z_2$ from the previous problem.
How does this plotted graph compare to the graph made in question 8?

```{r}
# insert code here
z1 <- (15 - 20) / 3
z2 <- (27 - 20) / 3

x <- seq(-4, 4, by = 0.1)
density <- dnorm(x)

df <- data.frame(x = x, density = density)

ggplot(df, aes(x = x, y = density)) +
  geom_line(color = "blue") +
  geom_ribbon(aes(ymin = 0, ymax = density, fill = (x > z1 & x < z2)), alpha = 0.5) +
  xlab("z") +
  ylab("Density") +
  ggtitle("Standard Normal Distribution (mean = 0, sd = 1)") +
  theme_minimal()
```



### 5

#### 5A

Using the chimpanzee data,
combine all of the trials made without a partner.
Count the number of pro-social and selfish choices made.

```{r}
# insert code here

```

#### 5B

Write a model for the number of pro-social choices.

#### 5C

Use the Agresti-Coull method to construct a 95% confidence interval for the probability that a chimpanzee makes the pro-social choice in experimental conditions when there is no partner present.

```{r}
# insert code here

```

#### 5D

Interpret this confidence interval in the context of the experimental setting.

> REPLACE THIS TEXT WITH YOUR RESPONSE



### 6

Using the same data as in the previous problem,
do the following.

#### 6A

State hypotheses for a two-sided hypothesis test that the probability that a chimpanzee makes the pro-social choice over the long run is equal to 0.5 versus the alternative that this probability is not 0.5.

> REPLACE THIS TEXT WITH YOUR RESPONSE

#### 6B

Pick a test statistic, state the sampling distribution of this statistic if the null hypothesis is true,
and compute a p-value for the test.

```{r}
# insert code here

```

#### 6C

Interpret these results in the context of the experiment following the examples from lecture notes.

> REPLACE THIS TEXT WITH YOUR RESPONSE



### 7

Use functions from the file `ggprob.R` to visualize the calculation of the p-value by graphing an appropriate probability distribution and using differential coloring to indicate what in the plot corresponds to the p-value.
Include a title and meaningful axis labels.

```{r}
# insert code here

```


